---
title: "Tuberculosis rates around the world for the last two decades"
author: "Astrid"
date: "2023-08-11"
always_allow_html: true
output:
  pdf_document:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```



```{r Import Packages, echo=TRUE, message=FALSE}
library(readr) # to import data
library(readxl) #to import excel file
library(dplyr) # to work with data
library(tidyr) # to change data formats
library(lubridate) #to work with date and time
library(magrittr) # to use pipes
library(ggplot2) # for graphs
library(plotly) # for visualisations
library(gapminder) # to import gdp data

```


```{r Import data set, echo=TRUE}
TBdata <- read_csv("../Data/TB and development data.csv")
Meta <- read_excel("../Data/Metadata.xlsx", 
    sheet = "Country - Metadata")
WHOTB <- read_csv("../Data/MDR_RR_TB_burden_estimates_2023-08-13.csv")

head(TBdata)
head(Meta)
head(WHOTB)
```

```{r tidy graph TB, echo=TRUE}
TBdata <- head(TBdata, -2) #remove bottom rows

TBdata <- TBdata[!is.na(TBdata$`Series Name`), ] #remove empty series names

names(TBdata) <- gsub("\\[.*?\\]", "", names(TBdata)) # remove [] from year columns

#change missing data to NA
TBdata[TBdata == '..'] <- NA

#Update column names
TBdata <- TBdata %>%
  mutate(`Series Name` = case_when(
    `Series Name` == "Incidence of tuberculosis (per 100,000 people)" ~ "TBincidence",
    `Series Name` == "Current health expenditure per capita (current US$)" ~ "Health_expenditure",
    `Series Name` == "GDP per capita (current US$)" ~ "GDPperCap",
    `Series Name` == "Life expectancy at birth, total (years)" ~ "Life_expectancy",
    `Series Name` == "Incidence of HIV, ages 15-49 (per 1,000 uninfected population ages 15-49)" ~ "HIVprevalence",
    TRUE ~ `Series Name` # Keep other values as they are
  ))

```






```{r Pivot Graph, echo=TRUE}
TBdata <- TBdata%>%
  rename(country= `Country Name`)

TBdata <- TBdata %>% #pivot longer
  pivot_longer(cols = 5:25, names_to = "year", values_to = "value")

TBdata <- TBdata %>% #pivot wider
  pivot_wider(names_from = `Series Name`, values_from = value,
              id_cols = c(country, year))


```

```{r Tidy metadata, echo=TRUE}
Meta <- Meta %>%
  select(`Table Name`, Region, `Income Group`, Code) %>%
  rename(
    country = `Table Name`,
    Income = `Income Group`
  )

TBdata <- inner_join(TBdata, Meta, by = "country")

```

```{r Fix up WHOTB data set, echo=TRUE}
WHOTB <- WHOTB %>%
  select(iso3, year, e_rr_pct_new) %>%
  rename(
    Code = iso3,
    ResisTB = e_rr_pct_new) 

# Convert year column in TBdata to numeric 
TBdata <- TBdata %>%
  mutate(year = as.numeric(year))

TBdata <- TBdata %>%
  mutate(TBincidence = as.numeric(TBincidence))

# Convert year column in WHOTB to numeric as well
WHOTB <- WHOTB %>%
  mutate(year = as.numeric(year))

WHOTB <- WHOTB %>%
  mutate(ResisTB = as.numeric(ResisTB))

# join
WHOTB <- WHOTB %>%
  inner_join(TBdata, by = c("Code", "year"))

WHOTB <- WHOTB %>%
  mutate(HIVprevalence = as.numeric(HIVprevalence))

```

```{r Fix up TB incidence, echo=TRUE}
#Find and delete data on NA if greater than 10 
TBdata <- TBdata %>%
  group_by(country) %>%
  mutate(NA_count = sum(is.na(TBincidence))) %>%
  filter(NA_count < 10) %>%
  select(-NA_count)

#change to numeric
TBdata <- TBdata %>%
  mutate(
    year = as.numeric(year),
    TBincidence = as.numeric(TBincidence),
    GDPperCap = round(as.numeric(GDPperCap), 2),
    Health_expenditure = round(as.numeric(Health_expenditure), 2),
    Life_expectancy = round(as.numeric(Life_expectancy), 2),
    HIVprevalence = as.numeric(HIVprevalence))

#Fill in missing data          
TBdata <- TBdata %>%
  mutate(year = as.numeric(year)) %>%
  group_by(country) %>%
  complete(year = min(year):max(year)) %>%
  fill(TBincidence, .direction = "downup")

sum(is.na(TBdata$TBincidence))

NATB <- TBdata[is.na(TBdata$TBincidence), "country"]

```

```{r Fix up Life expectancy, echo=TRUE}
NALE <- TBdata[is.na(TBdata$Life_expectancy), "country"]

TBLdata <- TBdata %>%
  group_by(country) %>%
  filter(!any(is.na(Life_expectancy))) %>%
  ungroup()


NAHIV <- TBdata[is.na(TBdata$HIVprevalence), "country"]

```

#### Tuberculosis Worldwide

```{r echo=TRUE, warning=FALSE}

plot_ly(data = TBLdata, ids = ~Code, frame = ~year) %>%
  add_trace(
    z = ~TBincidence,
    zmin = min(TBLdata$TBincidence), # set minimum value
    zmax = max(TBLdata$TBincidence), # set maximum value
    locations = ~Code,
    type = 'choropleth',
    colorscale = list(c(0, "white"), c(1, "darkblue"))
  ) %>%
  layout(
    title = "Global Tuberculosis Incidence Rates Over Time per 100,000 people",
    geo = list(
      showframe = FALSE,
      showcoastlines = FALSE,
      projection = list(type = 'equirectangular')
    ),
    autosize = TRUE
  ) %>%
  animation_opts(frame = 100, redraw = TRUE) %>%
  animation_slider(currentvalue = list(prefix = "Year: "))



```

Reference

The World Bank (2023) *Incidence of tuberculosis (per 100,000 people)*, The World Bank website, accessed 10 August 2023.
https://databank.worldbank.org/reports.aspx?dsid=2&series=SH.TBS.INCD




#### Tuberculosis vs Life Expectancy and HIV prevalence

```{r echo=TRUE, message=FALSE, warning=FALSE}
pTB2year <- plot_ly(data = TBLdata, x = ~Life_expectancy, y = ~TBincidence,
             text = ~country, 
             size = ~HIVprevalence, 
             color = ~Region, 
             frame = ~year,
             ids = ~country, alpha = 1) %>% 
  add_trace(type = "scatter", mode = "markers", 
            marker = list(sizemode = "diameter", sizemin = 3, sizeref = 3, 
              sizemax = 15)) %>% 
  layout(title = list(text = "Incidence of Tuberculosis by Year compared with Life Expentancy and HIV Prevalence World Wide", font = list(size = 12)), # 
         yaxis = list(
           zeroline = FALSE, 
           showgrid = FALSE, 
           title = list(text = "Incidence of Tuberculosis", font = list(size = 10)), 
           tickfont = list(size = 8) # Reduce y-axis tick label font size
         ),
         xaxis = list(
           zeroline = FALSE, 
           showgrid = FALSE, 
           title = list(text = "Life Expentancy", font = list(size = 10)), 
           tickfont = list(size = 8) # Reduce x-axis tick label font size
         ),
         showlegend = TRUE, 
         legend = list(title = list(text = 'Region (Size is prevalence of HIV)'),
                       font = list(size = 8)))



pTB2year
```

Reference

The World Bank (2023) *Incidence of tuberculosis (per 100,000 people)*, The World Bank website, accessed 10 August 2023.
https://databank.worldbank.org/reports.aspx?dsid=2&series=SH.TBS.INCD



#### New Rifampicin Resistance Infections in Europe 
```{r echo=TRUE, warning=FALSE}

WHOTB <- WHOTB %>%
  group_by(country) %>%
  filter((Region %in% c("Europe & Central Asia"))) %>%
  ungroup()

WHOTB <- WHOTB %>%
  group_by(country) %>%
  filter(!(Income %in% c("High income"))) %>%
  ungroup()



pWHO <- plot_ly(data = WHOTB, y = ~reorder(country, -TBincidence), frame = ~year) %>%
  add_trace(x = ~ResisTB, type = "bar", name = "Resistant Cases") %>%
    layout(
    title = list(text = "Percentage of New Rifamipicin Resistant Cases for Lower Income Countries in Europe and Central Asia", font = list(size = 12)),
    xaxis = list(
      title = list(text = "Percentage of Cases", font = list(size = 10)),
      showgrid = FALSE,
      tickfont = list(size = 8) 
    ),
    yaxis = list(
      title = list(text = "Country", font = list(size = 10)),
      title_standoff = 100,
      tickfont = list(size = 8) 
    ),
    barmode = 'stack',
    showlegend = TRUE,
    orientation = 'h',
    margin = list(l = 100), 
    legend = list(font = list(size = 6))
  ) %>%
  animation_opts(frame = 200, redraw = TRUE) %>% 
  animation_slider(currentvalue = list(prefix = "Year: "))






pWHO

```
Reference

World Health Organization (2022) *Tuberculosis data*, World Health Organization website, accessed 10 August 2023. https://www.who.int/teams/global-tuberculosis-programme/data


